An empirical, Bayesian approach to modelling the impact of weather on crop yield: maize in the US

### 2021-10-14 reviewer_response_2.tex

\begin{letter}{Environmental Research Communications \\ Institute of Physics \\ London, UK}

\opening{Dear editorial board,}

I am writing to respond to the reviewers' comments made following our submission of the paper ``An empirical, Bayesian approach to modelling the impact of weather on crop yield: maize in the US" to Environmental Research Communications.

This submission is made in response to the second set of comments made by the reviewers. Both reviewers comment that the re-submission made an improvement on the original but both have requested some further changes.

@@

As with the last submission we also provide an annotated version of the paper in addition to detailing changes that have been made here. The annotated version uses the following convention:
\paper{
Text that has not been changed will be rendered normally like this. \remove{We will highlight text that has been removed from the paper by showing it red and struck through like this.}\add{We will likewise show text that has been added by displaying it in green like this.}
}

We will now respond to the specific reviewer comments. The first reviewer begins:

\review{
REFEREE REPORT(S):
Referee: 1

COMMENTS TO THE AUTHOR(S)
The authors done a good job in revising the paper. I think the choice to switch towards Environmental Research Communications is a good choice. As I understand, novelty is no longer a criterion to consider.
}

We are grateful for the positive comment and agree that Environmental Research Communications is a more appropriate journal.

@@

\review{

In general, my points were addressed. Three points remain:
- I still struggle with the comparison to Schlenker and co-authors. You now highlight several times that you provide a ‘simple and robust model’. Please clarify: that i) that the here presented approach is still scientific valid – and where are the limits, ii why is the Schlenker et al approach not ‘simple and robust’?; iii) what is the value added from a scientific and practical point of view?

}

The simplification is the movement from daily maximum to monthly means. This is done to increase the applicability to data sets that only contain monthly means. The robustness is due to the Bayesian framework which properly characterises the posterior to give a measure of the uncertainty in each model parameter given the data and the model and the uncertainty in predicted yields. Crucially, it incorporates model assumptions into uncertainties. We have removed some of the instances of this phrase and explained it at the first usage. The value added is that predictions of the impact of global changes to temperature and precipitation combine the uncertainties in those changes with uncertainties in the model to give a measure of the uncertainty in the predicted yields.

Here, we have access to an average increase of temperature and precipitation across the whole year from climate simulations. It would therefore not add predictive power to resolve at the daily frequency. We are targeting the general impacts of climate change to yields.

@@

\review{
- Section 2 is now data. On lines 151 ff you sketch possible impacts of climatic extremes. Given the relevance of these fundamentals, I invite you to add more details and literature. I think your paper benefits from having a sting agronomic foundation why and how heat and drought matter for the empirical analysis to come

}

We have added additional paragraphs to further discuss the literature.

\review{
- Line 130ff: I am still puzzled with the detrending vs median. Conceptually, if yields are continuously increasing over time, a median based approach will underestimate the yield residuals (because the yield is always higher than the median from previous years because of technological development). Whatever your comparisons show, I think this is a bias that matters, especially if you want to identify ‘extremes’. Given the relevance of extreme yield observations and climatic extremes, a robust detrending might be suitable (see e.g. Finger 2010). Some additional mode runs with a detrending might be added as appendix
}

The five year rolling median is based on the previous two years, the current year, and the next two years. The median bias will actually only impact the two first and two last values. Away from the boundary there is no bias (31 out of 35 years for each state). In a predictive framework where you only have previous years the anomaly is therefore only with respect to the as yet unknown median. In this case the linear detrending is the only method that does not suffer from this effect. However, the model here can still be used to predict the climactic impact on yields as distinct from the impact of increased productivity. In particular this will be useful in extremes of climate that will have a significant impact on the yield.

Finally we recommend using fractional anomalies from the five year rolling mean. As with the median the bias will only impact the edge cases and so will have a small impact on the final parameter values. The Finger 201 detrending method is a robust means (insensitive to outliers) to compute estimators for a detrending model. It could be applied for a linear productivity increase model in order to generate new linearly detrended anomalies. However, figure 1 shows that a least squares estimator captures the productivity increase well. Here the model performs best on the mean five year anomalies which would not be improved using the Finger 2010 method of moments estimator technique since there are no parameters to be estimated. As a test we tried removing outliers with a fractional anomaly above ten percent and found that is had a low impact on the parameters of the linear fit and a negligible impact (changes to median values of less than 5~\% of the variance) on the posterior parameter values derived from those anomalies.

\review{
- Discussion: I still think you shall critically discuss that your focus on months is not appropriate to capture sensitive crop yield growth stages because these vary from year to year. Instead, a consideration of phenology can help to overcome this

@@

Finger, R. (2010). Revisiting the Evaluation of Robust Regression Techniques for Crop Yield Data Detrending. American Journal of Agricultural Economics 92(1): 205-211
}

The model's insensitivity to daily temperature changes will become important with access to detailed climate simulations which will predict the impact of a changing climate on both mean increases and changes to variability on a daily scale. However, we do not have access to such high resolution predictions and so the challenge we aim to address here is to determine the predictive power of month resolution modelling and the impact of literature predictions of climate change. The model can account for key growth periods occurring in different months as it responds independently to each month's climate conditions.

@@

https://iopscience.iop.org/article/10.1088/1748-9326/ab422b/meta
}

We thank the reviewer for pointing us to this paper. The results are broadly in agreement with our results and it adds a further reference to compare the impact of temperature changes and precipitation changes. We have included it in the introduction.

\review{
Negative Classical $R^2$ in table 2? This means that the model preforms worse than the mean of the observed values. Assuming this was computed correctly... $R^2$ is not used for model predictive ability. For that you have, for example, the Mean Squared Error of Prediction (MSEP) or Cross-Validation Error. I can’t find Table 2 referenced anywhere in the manuscript.
}

The output of a Bayesian model for each predicted yield is a probability distribution. The classical $R^2$ requires a point prediction in order to be computed. In order to do so here, we use the median posterior parameter values to compute the prediction. However, this is somewhat arbitrary as discussed in Gelman et al. (2019) since there are various choices that could be made (the median predicted value from the full sample of parameter values could be another choice). Without using any of the validation schemes it is true that classical $R^2$ compares these point prediction performance to the mean. However, using the mean can produce a negative value under the validation schemes which remove the target from the training set to avoid over-fitting. There is extensive literature on the interpretation of $R^2$ and we have used the Bayesian $R^2$ metric in its place following the discussion in Gelman et al. (2019) which highlights the possibility for the classical $R^2$ to lie outside the [0,1] interval under Bayesian inference approaches. It highlights situations where maximum likelihood (the equivalent of taking the posterior mode to make point predictions) can mask issues with model performance when using classical $R^2$. All these difficulties with classical $R^2$ are especially relevant in situations such as this where measurements are subject to large noise (measurement error and other factors comparable to the variance of interest). The table was not referenced in the previous submission due to a latex error which has been corrected.

### 2021-10-14 main.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams} 
\usepackage{xcolor}
\usepackage{newtxtext, newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\DeclareGraphicsExtensions{ .pdf, .png}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
%\usepackage{breqn}
\usepackage{hyperref}

\usepackage{lineno}
\linenumbers

%uncomment appropriate defs for annotated/not annotated copy
%\newcommand{\remove}[1]{
%    {\color{red} \sout{#1}}
%    }

\newcommand{\remove}[1]{}

%\newcommand{\add}[1]{   
%    {\color{green} #1}
%    }
\newcommand{\add}[1]{#1}

% Don't change these lines
\begin{document}
\label{firstpage}
%\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\title[Bayesian crop yield prediction]{An empirical, Bayesian approach to modelling the impact of weather on crop yield: maize in the US}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Shirley et al.]{
%\parbox{\linewidth}{
Raphael Shirley,$^{1,2,3}$ %\thanks{E-mail: raphael.shirley@sussex.ac.uk }
Edward Pope,$^{4}$
Myles Bartlett,$^{5}$
Seb Oliver,$^{1,2}$
Novi Quadrianto,$^{5}$
Peter Hurley,$^{6}$
Steven Duivenvoorden,$^{1,2}$
Phil Rooney,$^{6}$
Adam B. Barrett,$^{2}$
Chris Kent,$^{4}$
and James Bacon.$^{4}$
%}
}

\\
\\
% List of institutions
\parbox{\linewidth}{
{\scriptsize
$^{1}$Astronomy Centre, Department of Mathematical and Physical Sciences, University of Sussex, UK\\
$^{2}$Data Intensive Science Centre, Department of Mathematical and Physical Sciences, University of Sussex, UK\\
$^{3}$Instituto de Astrof\'{i}sica de Canarias, E-38205 La Laguna, Tenerife, Spain;\\ Universidad de La Laguna, Dpto. Astrof\'{i}sica, E-38206 La Laguna, Tenerife, Spain\\
$^{4}$Met Office Hadley Centre, Exeter, UK\\
$^{5}$Predictive Analytics Lab, Department of Engineering and Informatics, University of Sussex, UK\\
$^{6}$Data Javelin Ltd, Data science consultancy, Brighton, UK
}
}
% These dates will be filled out by the publisher
%\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
%\pubyear{2019}

% Abstract of the paper
%\section*{General comments}
%{\color{green}Seb: I think we need some consistency around using ``yield" or ``yields".  I guess if we are talking about one crop it should be yield and many crops yields?  But then is our model generic for multiple crops?  Alternatively one could decide that it is singular throughout?}
%Raph: I went for using singular in general

\begin{abstract}
We apply an empirical, data-driven approach for describing crop yield as a function of monthly temperature and precipitation by employing generative probabilistic models with parameters determined through Bayesian inference. Our approach is applied to state-scale maize yield and meteorological data for the US Corn Belt from 1981 to 2014 as an exemplar, but would be readily transferable to other crops, locations and spatial scales. Experimentation with a number of models shows that maize growth rates can be characterised by a two-dimensional Gaussian function of temperature and precipitation \add{with monthly contributions accumulated over the growing period. This approach accounts for non-linear growth responses to the individual meteorological variables, and allows for interactions between them}. Our models correctly identify that temperature and precipitation have the largest impact on yield in the six months prior to the harvest, in agreement with the typical growing season for US maize (April to September). Maximal growth rates occur for monthly mean temperature 18-19$^\circ$C, corresponding to a daily maximum temperature of 24-25$^\circ$C (in broad agreement with previous work) and monthly total precipitation 115 mm. Our approach also \add{provides}\remove{facilitates} a self-consistent \add{way of} \remove{investigation of}\add{investigating} climate change impacts on \add{current} US maize \add{varieties}\remove{yield} in the absence of adaptation measures. Keeping precipitation and growing area fixed, a temperature increase of $2^\circ$C, relative to 1981-2014, results in the mean yield decreasing by 8\%, while the yield variance increases by a factor of around 3.  We thus provide a flexible, data-driven framework for exploring the impacts of natural climate variability and climate change on globally significant crops based on their observed behaviour. In concert with other approaches, this can help inform the development of adaptation strategies that will ensure food security under a changing climate.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
Bayesian inference -- crop yield -- weather -- climate
\end{keywords}

@@

Establishing the climate risk to the global production of individual crops, and how that might change in the future, is an essential requirement for building a resilient and robust food system that ensures food security for all \citep[][]{fao:2002}. %However, quantifying this risk represents a significant challenge, primarily because limited data records hinder our ability to identify causal 
This requires the characterisation of the relationship between meteorological and food production variations.

There is a growing consensus that a range of methods are needed to accurately assess climate impacts on crop yield \citep[e.g.][]{lobell:2017, tigchelaar:2018, snyder:2018}.  Broadly there are two complementary approaches to this:  physiological processes-based models; or data-driven, statistical models.

Data-driven, statistical methods can be rapidly applied when there is yield data, even where there is limited laboratory-based understanding of the crop. Physiological models, however, can be extrapolated into situations where there is limited yield data.

Previous statistical methods \citep[e.g.][]{schlenker:2009, lobell:2010, welch:2010} capture the broad influence of weather on crop yield based on linear relationships, calibrated using historical data. Non-linear models have been explored based on temperature \citep[e.g.][]{cutforth:1990, yin:1995, wang:1998, yan:1999, streck:2007, zhou:2018} and precipitation  \citep[e.g.][]{cakir:2004, ge:2012, lobell:2013, carter:2016, song:2019}.

By deriving an empirical \add{and non-linear} ``growth response function'' that \remove{maps}\add{relates} temperature and rainfall conditions to crop growth, the approach developed here aims to complement both process-based crop models and existing statistical approaches.

%For this study, we use monthly (rather than daily) data for two reasons: firstly, to extract the typical crop response to large-scale meteorological conditions; secondly, to test the performance of the model with limited data as a way of assessing its potential utility in regions where historical weather records may be less complete.

%\add{To date,} the viability of a data-driven approach for describing maize yield as a function of monthly temperature and precipitation has not been demonstrated. \add{However,} this capability could would allow us to gain \add{new} insight into the potential implications for maize yield under a changing climate since \remove{current simulations do not capture}\add{we have more confidence in projected changes on monthly rather than daily timescales}\remove{daily climatalogical variation}.

%{Ed - Trying to motivate our approach a bit more, and explaining why we are testing the procedure for US maize, and that this can provide info for crop breeders and agronomists to target specific developments}.

%{\color{green} [not sure if this sentence belongs in the introduction and, if so, if this is the right place]} {Crucially, the output model parameters can be compared to known

%{\color{green} this para needs to be a bit clearer, particularly the second sentence.  I guess the point being made is that different methods are complementary, but this message seems slightly implicit rather than explicit - and in fact it is hard to pick out exactly what is being said}
%Raph - added note that absence of impact of CO2 is main drawback of empirical models compared to process based models (taken from lobell:2017)

This work is part of a trend to apply advanced statistics and machine learning methods to climatological and agricultural data sets. For example, \citet{you:2017} demonstrated the application of deep learning and Gaussian processes to predict yield based on remote imaging.

Bayesian inference, is an established approach for inferring the posterior values of model parameters, based on prior assumptions and new data. The advantage of Bayesian inference is it allows robust computation of errors, which is especially critical when the aim is to model the influence of predicted climate data which are themselves subject to large uncertainties.

More generally, Bayesian inference is frequently applied to problems of model parameter estimation with noisy data in other fields such as astronomy \citep{Hurley:2016}. In this paper, we are determining generative probabilistic models, which have a greater ability to accurately capture uncertainty than the more common discriminative models in machine learning.

The aim here is to present a simple and robust model which captures the impact of mean temperature and precipitation changes on mean yield. Developing a flexible method for investigating the influence of present-day natural climate on yield allows us to explore the direction and scale of climate change impacts in the absence of adaptation.

Here, we use these methods to model the response of US maize to temperature and precipitation, which has been widely studied \citep[e.g.][]{schlenker:2009, hatfield:2011, roberts:2012, lobell:2013, sanchez:2014, hatfield:2015, Partridge:2019}. We aim to first demonstrate an ability to capture key aspects of present-day maize yield variability in the US, and secondly to explore the implications of climate change for the current maize varieties in the absence of adaptation.

The paper is structured as follows: section~\ref{sec:data} describes the yield and meteorological data, followed by section~\ref{sec:models} which presents the models, demonstrating their strengths and weaknesses. Section~\ref{sec:impacts} applies the model \remove{with increments in temperature and precipitation}\add{to an ensemble of climate projections} as a first step to predict the \remove{consequence}\add{influence} of climate change on \remove{agricultural}yield \add{for present day maize varieties}. Finally, section~\ref{sec:conclusions} summarises the work and our main conclusions.

@@

\add{Numerous studies have demonstrated the importance of temperature in maize development \citep[e.g.][]{cross:1972, coelho:1980, daughtry:1984, cutforth:1990, bonhomme:1994}, with quantities such as Growing Degree Days (GDD) offering better predictions of phenological changes and yield than calendar days after planting. Water stress is also associated maize yield reductions \citep[e.g.][and references therein]{cakir:2004, ge:2012, lobell:2013, carter:2016, song:2019}, suggesting that any model should incorporate the effects of both temperature and precipitation.}

We \add{proceed} assuming that climate variability is \add{a major}\remove{the dominant} driver of the \remove{resulting}\add{observed} yield anomalies, and do not attempt to separate direct physiological influences from non-direct impacts resulting from air quality or pests \citep{Gornall:2010}.

We use annual maize yield and monthly temperature and rainfall data aggregated to the state scale. This minimises the impact of local variations in both meteorology and planting date \citep[e.g.][]{schlenker:2009, lobell:2010}. For future work, the model outputs are compatible with the large-scale atmospheric circulation patterns that can be reliably simulated by global climate models. The relatively low data requirements, compared to daily climatology, also support model flexibility and computationally efficiency. Constraining the model parameter posteriors depends on the sampling of temperature precipitation space, the data can only constrain features which are present in the data. We therefore draw observations from across the US Corn Belt, covering Indiana, Illinois, Iowa, Ohio, Minnesota and Nebraska to maximise the sampling while being confident the regions use similar agricultural techniques and have broadly comparable climatology.

@@

In all regions there has been a long-term increase in maize yield since 1960, on top of which year-to-year and multi-year variations are evident. To remove the long-term trend and decadal-scale variability, which can be driven by climate and non-climate factors \citep[e.g.][]{hawkins:2013, ray:2015} we use yield `anomalies'. For median yield anomalies the centred 5-year rolling median yield in each region is subtracted from the time series. The remaining yield fluctuations retain the units of tonnes per hectare (t/ha). We also compare \remove{anolies}\add{anomalies} from the 5-year mean, fractional anomalies from 5-year mean (which are unit-less), and linearly detrended anomalies from the least squares linear trend across all regions. \add{\citet{finger:2010}, applied robust detrending techniques, which are designed to perform regression in the presence of outliers. Such an approach here could be used to perform the linear detrending, however it would still rely on the assumption of a linear productivity increase.}

The total yield anomaly in year $j$, $\Delta Y_j$, is the difference between the yield for year $j$, $Y_j$, and the rolling five-year median or mean $\widetilde{Y}_j = (Y_{j-2},Y_{j-1},Y_{j},Y_{j+1},Y_{j+2} )$. A feature of using the rolling 5-year median anomalies is that, on average, one in five anomalies will be exactly zero. For this reason we compare all four forms of yield anomaly. The choice of de-trending has some influence on the posterior parameter values and, therefore, the model performance as described in section~\ref{sec:validation}. The temperature and precipitation anomalies for each month are calculated by subtracting the corresponding monthly climatological mean for 1981-2014.

The relationship between the yield and the temperature and precipitation anomalies forms the basis of our linear models. These provide a useful comparator for more complex models that relate yield to temperature and precipitation. Table~\ref{table:data} summarises the variation around typical values for monthly temperatures and precipitation and annual maize yields. Figure~\ref{fig:yield_anomalies} illustrates how these two measures are related, while Figure~\ref{fig:median_vs_frac_anoms_noline} shows the relation between the 5-year median anomaly and the 5-year fractional mean anomaly. Later, we compare what impact the choice of target yield has on the performance of models.

@@

Previous analysis of these data by \cite{kent:2017} revealed that yield reductions greater than 10\% are strongly associated with mean temperatures during June, July and August exceeding 23 $^\circ$C, combined with total precipitation less than 240 mm. In short, warmer and drier than normal conditions were associated with reduced maize yield in the US Cornbelt. There was also evidence that excess precipitation during the same period could predict yield reductions; however, this relationship was more tentative. %{\color{green} The following sentence is more about our method than the data - should this go somewhere else?} The approach developed here aims to explore the context for these results through providing a more general and continuous description of the crop's response to temperature and precipitation, while building on the threshold-based approach demonstrated in \cite{kent:2017}. 
% Raph - moved it to the description of 2d model in models

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%use iopart possibly but I can only remove all errors using article
\documentclass[12pt]{article}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams} 
\usepackage{xcolor}
%\usepackage{newtxtext, newtxmath}
%\usepackage{amsmath}	% Advanced maths commands
%\usepackage{amssymb}	% Extra maths symbols

% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\DeclareGraphicsExtensions{ .pdf, .png}

%\usepackage{breqn}
\usepackage{hyperref}

% Don't change these lines
\begin{document}
%\label{firstpage}
%\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\title{An empirical, Bayesian approach to modelling the impact of weather on crop yield: maize in the US}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author{
\parbox{\linewidth}{
Raphael Shirley,$^{1,2,3}$ %\thanks{E-mail: raphael.shirley@sussex.ac.uk }
Edward Pope,$^{4}$
Myles Bartlett,$^{5}$
Seb Oliver,$^{1,2}$
Novi Quadrianto,$^{5}$
Peter Hurley,$^{6}$
Steven Duivenvoorden,$^{1,2}$
Phil Rooney,$^{6}$
Adam B. Barrett,$^{2}$
Chris Kent,$^{4}$
and James Bacon.$^{4}$
}
}
\maketitle

% List of institutions
\parbox{\linewidth}{
{\scriptsize
$^{1}$Astronomy Centre, Department of Mathematical and Physical Sciences, University of Sussex, UK\\
$^{2}$Data Intensive Science Centre, Department of Mathematical and Physical Sciences, University of Sussex, UK\\
$^{3}$Instituto de Astrof\'{i}sica de Canarias, E-38205 La Laguna, Tenerife, Spain;\\ Universidad de La Laguna, Dpto. Astrof\'{i}sica, E-38206 La Laguna, Tenerife, Spain\\
$^{4}$Met Office Hadley Centre, Exeter, UK\\
$^{5}$Predictive Analytics Lab, Department of Engineering and Informatics, University of Sussex, UK\\
$^{6}$Data Javelin Ltd, Data science consultancy, Brighton, UK
}
}

\begin{abstract}
We apply an empirical, data-driven approach for describing crop yield as a function of monthly temperature and precipitation by employing generative probabilistic models with parameters determined through Bayesian inference. Our approach is applied to state-scale maize yield and meteorological data for the US Corn Belt from 1981 to 2014 as an exemplar, but would be readily transferable to other crops, locations and spatial scales. Experimentation with a number of models shows that maize growth rates can be characterised by a two-dimensional Gaussian function of temperature and precipitation with monthly contributions accumulated over the growing period. This approach accounts for non-linear growth responses to the individual meteorological variables, and allows for interactions between them. Our models correctly identify that temperature and precipitation have the largest impact on yield in the six months prior to the harvest, in agreement with the typical growing season for US maize (April to September). Maximal growth rates occur for monthly mean temperature 18-19$^\circ$C, corresponding to a daily maximum temperature of 24-25$^\circ$C (in broad agreement with previous work) and monthly total precipitation 115 mm. Our approach also provides a self-consistent way of investigating climate change impacts on current US maize varieties in the absence of adaptation measures. Keeping precipitation and growing area fixed, a temperature increase of $2^\circ$C, relative to 1981-2014, results in the mean yield decreasing by 8\%, while the yield variance increases by a factor of around 3.  We thus provide a flexible, data-driven framework for exploring the impacts of natural climate variability and climate change on globally significant crops based on their observed behaviour. In concert with other approaches, this can help inform the development of adaptation strategies that will ensure food security under a changing climate.
\end{abstract}

\noindent{\it Keywords\/}: Bayesian inference, crop yield, weather, climate

@@

Establishing the climate risk to the global production of individual crops, and how that might change in the future, is an essential requirement for building a resilient and robust food system that ensures food security for all \citep[][]{fao:2002}. Decision-makers can then use this information to guide the development of suitable adaptation and mitigation strategies across different time frames. This requires the characterisation of the relationship between meteorological and food production variations.

There is a growing consensus that a range of methods are needed to accurately assess climate impacts on crop yield \citep[e.g.][]{lobell:2017, tigchelaar:2018, snyder:2018}. Exploring different model formulations and assumptions (e.g. a multi-model ensemble) provides a way of assessing key uncertainties and biases in our understanding of crop-climate interactions. In turn, this can help evaluate our confidence in the direction and magnitude of climate change impacts on food production. Broadly, there are two complementary approaches to this:  physiological processes-based models; or data-driven, statistical models. Physiological models are generally built upon an experiment-based understanding of the generic crop. Data-driven, statistical models can be developed when there is sufficient empirical yield data. Each approach can thus be developed and applied when the requirements for the other are not met.

Examples of statistical approaches include non-parametric models which are formulated in terms of meteorological variables rather than underlying physiological processes or critical thresholds. These models are calibrated using historical data and have demonstrated the ability to capture broad influences of weather on crop yield \citep[e.g.][]{schlenker:2009, lobell:2010, welch:2010}. This approach differs to parameterised models, calibrated by field experiments, which do account for specific thresholds in quantifying the response to temperature \citep[e.g.][]{cutforth:1990, yin:1995, wang:1998, yan:1999, streck:2007, zhou:2018} or precipitation \citep[e.g.][]{cakir:2004, ge:2012, lobell:2013, carter:2016, song:2019}. Observations have also been used to constrain parameters of more complex process-based models \citep[e.g.][]{iizumi:2009, tao:2009}. A physically motivated, but empirical, data-driven approach would complement both process-based crop models and existing statistical approaches. This formulation would allow models to be developed without extensive field trials, and with a greater range of validity.

Our approach exploits Bayesian inference to derive an empirical and non-linear ``growth response function'' that maps temperature and rainfall conditions to crop growth. This work is part of a trend to apply advanced statistics and machine learning methods to climatological and agricultural data sets. For example, \citet{you:2017} demonstrated the application of deep learning and Gaussian processes to predict yield based on remote imaging.

Bayesian inference, is an established approach for inferring the posterior values of model parameters, based on prior assumptions and new data. The advantage of Bayesian inference is it allows robust computation of errors, which is especially critical when the aim is to model the influence of predicted climate data which are themselves subject to large uncertainties. More generally, Bayesian inference is frequently applied to problems of model parameter estimation with noisy data in other fields such as astronomy \citep{Hurley:2016}. In this paper, we are determining generative probabilistic models, which have a greater ability to accurately capture uncertainty than the more common discriminative models in machine learning.

The aim here is to present a simple and robust model which captures the impact of mean temperature and precipitation changes on mean yield. Developing a flexible method for investigating the influence of present-day natural climate on yield allows us to explore the direction and scale of climate change impacts in the absence of adaptation. We apply these methods to model the response of US maize to temperature and precipitation, which has been widely studied \citep[e.g.][]{schlenker:2009, hatfield:2011, roberts:2012, lobell:2013, sanchez:2014, hatfield:2015, Partridge:2019}. We aim to first demonstrate an ability to capture key aspects of present-day maize yield variability in the US, and secondly to explore the implications of climate change for the current maize varieties in the absence of adaptation.

The paper is structured as follows: section~\ref{sec:data} describes the yield and meteorological data, followed by section~\ref{sec:models} which presents the models, demonstrating their strengths and weaknesses. Section~\ref{sec:impacts} applies the model to an ensemble of climate projections as a first step to predict the influence of climate change on yield for present day maize varieties. Finally, section~\ref{sec:conclusions} summarises the work and our main conclusions.

@@

Numerous studies have demonstrated the importance of temperature in maize development \citep[e.g.][]{cross:1972, coelho:1980, daughtry:1984, cutforth:1990, bonhomme:1994}, with quantities such as Growing Degree Days (GDD) offering better predictions of phenological changes and yield than calendar days after planting. Water stress is also associated maize yield reductions \citep[e.g.][and references therein]{cakir:2004, ge:2012, lobell:2013, carter:2016, song:2019}, suggesting that any model should incorporate the effects of both temperature and precipitation.

We proceed assuming that climate variability is a major driver of observed yield anomalies, and do not attempt to separate direct physiological influences from impacts resulting from air quality or pests \citep{Gornall:2010}. We use annual maize yield and monthly temperature and rainfall data aggregated to the state scale. This minimises the impact of local variations in both meteorology and planting date \citep[e.g.][]{schlenker:2009, lobell:2010}. For future work, the model outputs are compatible with the large-scale atmospheric circulation patterns that can be reliably simulated by global climate models. The relatively low data requirements, compared to daily data, also support model flexibility and computationally efficiency. Constraining the model parameter posteriors depends on the sampling of temperature/precipitation space, the data can only constrain features which are present in the data. We therefore draw observations from across the US Corn Belt, covering Indiana, Illinois, Iowa, Ohio, Minnesota and Nebraska to maximise the sampling while being confident the regions use broadly similar agricultural techniques and have comparable climatologies.

@@

In all regions there has been a long-term increase in maize yield since 1960, on top of which year-to-year and multi-year variations are evident. To remove the long-term trend and decadal-scale variability, which can be driven by climate and non-climate factors \citep[e.g.][]{hawkins:2013, ray:2015}, we use yield `anomalies'. Removing the estimated long term trend gives an estimate of the yield anomaly relative to what was expected that year, which retains the units of tonnes per hectare (t/ha). We compare a range of methods for calculating yield anomalies: a) anomalies relative to the centred 5-year rolling median yield in each region which is subtracted from the time series; b) anomalies relative to the 5-year rolling mean; c) fractional anomalies relative to the 5-year rolling mean (which are unit-less); d) anomalies relative to the least squares linear trend across all regions. \citet{finger:2010}, applied robust detrending techniques, which are designed to perform regression in the presence of outliers. Such an approach could be used here to perform the linear de-trending; however, it would still rely on the assumption of a linear productivity increase. Removing outliers beyond 10\% fractional yield anomaly influenced the linear fit parameters by less than 10\% and had a negligible impact on parameter posteriors (changes to median values of less than 5~\% of the variance).

The total yield anomaly in year $j$, $\Delta Y_j$, is the difference between the yield for year $j$, $Y_j$, and the rolling five-year median or mean $\widetilde{Y}_j = (Y_{j-2},Y_{j-1},Y_{j},Y_{j+1},Y_{j+2} )$. A feature of using the rolling 5-year median anomalies is that, on average, one in five anomalies will be exactly zero. For this reason we compare all four forms of yield anomaly. The choice of de-trending has some influence on the posterior parameter values and, therefore, the model performance as described in section~\ref{sec:validation}. The temperature and precipitation anomalies for each month are calculated by subtracting the corresponding monthly climatological mean for 1981-2014. However, the main model presented uses actual temperature and precipitation values and not anomalies.

The relationship between the yield and the temperature and precipitation anomalies forms the basis of our linear models. These provide a useful comparator for more complex models that relate yield to temperature and precipitation directly. Table~\ref{table:data} summarises the variation around typical values for monthly temperatures and precipitation and annual maize yields. Figure~\ref{fig:yield_anomalies} illustrates how these two measures are related, while Figure~\ref{fig:median_vs_frac_anoms_noline} shows the relation between the 5-year median anomaly and the 5-year fractional mean anomaly. Later, we compare what impact the choice of target yield has on the performance of models.

@@

Previous analysis of these data by \cite{kent:2017} revealed that yield reductions greater than 10\% are strongly associated with mean temperatures during June, July and August exceeding 23 $^\circ$C, combined with total precipitation less than 240 mm. In short, warmer and drier than normal conditions were associated with reduced maize yield in the US Cornbelt. There was also evidence that excess precipitation during the same period could predict yield reductions; however, this relationship was more tentative.

Figure~\ref{fig:real_yield_responses} shows the maize yield time series, normalised to 2007 linear trend levels, as a function of yearly average temperature and precipitation. 
Even at this level of temporal coarse-graining, (i.e. averaging temperature and precipitation anomalies over six months), there is evidence that warmer and drier conditions are associated with significantly reduced maize yield, and the basic features that any model should try to capture are evident.
